Political Polarization, Public Trust, and Healthcare Disparities: Unraveling COVID-19's Impact and Future Preparedness

### 2024-02-14 outputs/paper/paper.qmd
Updated abstract

---
title: "Political Polarization, Public Trust, and Healthcare Disparities: Unraveling COVID-19's Impact and Future Preparedness"
author: 
  - Adrian Ly
  - Hannah Yu
  - Sakhil Goel
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git 
This reproduction was performed after a replication on the Social Science Reproduction platform: [link here](https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index)"
date: today
date-format: long
abstract: "The COVID-19 pandemic has exposed deep disparities in health outcomes across different states and counties in the United States. We use data from the American Economic Association to conduct a comprehensive analysis of how variances in local health infrastructure and socio-political factors within the United States contributed to disparities in COVID-19 outcomes. Our analysis reveals that political polarization significantly influenced adherence to health guidelines, with Republican-leaning states demonstrating slightly higher clustering around elevated death rates. Moreover, social vulnerabilities, including disparities in healthcare access, housing conditions, and ethnic demographics, emerged as key contributors to substantial variations in disease outcomes. These results highlight the urgent requirement for customized public health readiness and response tactics, focusing on precise interventions to tackle deficiencies in community health infrastructure and socio-political factors."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

Our analysis is grounded in comprehensive data sourced from the American Economic Association, enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes. While leveraging the initial findings from the AEA paper, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities. We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes. One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.

>>>>

---
title: "Political Polarization, Public Trust, and Healthcare Disparities: Unraveling COVID-19's Impact and Future Preparedness"
author: 
  - Adrian Ly
  - Hannah Yu
  - Sakhil Goel
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git 
This reproduction was performed after a replication on the Social Science Reproduction platform: [link here](https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index)"
date: today
date-format: long
abstract: "The COVID-19 pandemic has exposed deep disparities in health outcomes across various regions in the United States. Using research conducted by @citearticle, we analyze how variances in local health infrastructure and socio-political factors influenced COVID-19 outcomes. Our analysis reveals that political polarization notably impacted adherence to health guidelines, particularly in Republican-leaning states, correlating with higher death rates. Furthermore, disparities in healthcare access and housing significantly affected the life expectancy of different ethnic groups. These findings emphasize the need for tailored public health strategies to address infrastructure and socio-political deficiencies, while future research should continue to explore these factors for better pandemic preparedness and response."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

Our analysis is grounded in comprehensive data sourced from the paper, `Why Did the Best Prepared Country in
the World Fare So Poorly during COVID?` [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes. While leveraging the initial findings from the @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities. We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes. One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.
